---
title: "01_code_laboratory_analysis_Kettlitz et al."
author: "Robyn Kettlitz"
date: "2024-11-17"
output: html_document
---
R code of the article: 
Kettlitz et al., Feasibility and diagnostic performance of blood self-sampling with HemaSpot HF: The use case of Anti-Clostridium tetani toxin IgG detection.

This script was used to analyse the the diagnostic performance of HemaSpot HF device (HS) [1] in the detection of Anti-Clostridium tetani toxin IgG as a use case for comparison with  venous blood puncture (VBP) as the gold standard for blood collection. 

For a description of variables please refer to the codebook.

Abbreviation: 

HS = HemaSpot
HF = HemaForm
VBP = venous blood puncture

References: 

[1]: Sciences SO. HemaSpot™-HF Blood Collection Device  [Available from: https://www.spotonsciences.com/hemaspot-hf/ ]
[2]: NovaTec’s Clostridium tetani toxin 5S IgG plus – ELISA (PTETG043, NovaTec, Germany)

Use of artificial intelligence:
We used OpenAI's ChatGPT v2 to review and refine the R code for improved structure and functionality 


Code review by Ruibing Shi [Department for Epidemiology, Helmholtz Centre for Infection Research, Braunschweig, Lower Saxony, Germany]
-----------------------------
Codebook:

Variable name| Content question| Answer options

tet_hs | Result U/ml tetanus HS | continuous
tet_v | Result U/ml tetanus VBP | continuous
tet_p | Result U/ml tetanus HF  | continuous

hs1 | Recieved HS device | 0: no, 1: yes
auf0 | HS without conspicuous features | 0: no, 1: yes
auf1 | HS with external conspicuous features | 0: no, 1: yes
auf2 | HS with internal conspicuous features | 0: no, 1: yes
besch1 | HS external damage | 0: no, 1: yes

-----------------------------

##00: Set up, loading relevant packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(pacman)

pacman::p_load(tidyverse, gfonts ,haven, epiR,irr, BlandAltmanLeh,psych, Kendall, DescTools)
```

##01: Read data
```{r}
#identify newest file
recent_file <- list.files(path = "../../../5_Daten/Labordaten/Bearbeitet/", pattern = "MergedAndCleanedData_jor", full.names = TRUE)  %>% sort(decreasing = TRUE) %>% 
  .[1]

hema <- read_csv(recent_file) %>% 
  select(-1)
```

##02: Frequency of HS, HF, and VBP plus internal and external conspicuities 
```{r}

#Amount of HS and VBP samples
hema %>% filter(!is.na(tet_hs)) %>% count()
hema %>% filter(!is.na(tet_hs), is.na(tet_v)) %>% count()#HS without VBP
hema %>% filter(!is.na(tet_hs), !is.na(tet_p)) %>% count()#HS and HF
hema %>% filter(!is.na(tet_hs), !is.na(tet_v)) %>% count()#HS plus VBP


#Percentage: Received HS 
hema %>% 
  group_by(hs1) %>% 
  summarise(count = n()) %>%
  drop_na() %>% 
  mutate(percentage = round(count / sum(count)*100,digits=1))

#Percentage of HS with conspicuities
hema %>% 
  group_by(auf0) %>% 
  summarise(count = n()) %>%
  drop_na() %>% 
  mutate(percentage = round(count / sum(count)*100,digits=1))

#Percentage of HS with conspicuous external
hema %>% filter(auf0==0) %>% 
  group_by(auf1) %>% 
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count)*100,digits=1))


#Percentage of HS with conspicuous internal
hema %>% filter(auf0==0) %>% 
  group_by(auf2) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = round(count / sum(count)*100,digits=1))
  
```
##03:Preparation of the analysed labortaory values for detection in Anti-Clostridium tetani toxin IgG 
according to the manufacture manual of the ELISA (PTETG043, NovaTec, Germany) [2]
```{r}
#Store the laboratory values of the antibody concentration in the samples HS, HF, and VBP as numeric variables for analysis
hema$tet_hs<-as.numeric(hema$tet_hs)
hema$tet_v<-as.numeric(hema$tet_v)
hema$tet_p<-as.numeric(hema$tet_p)

#create binary variables for HemaSpot results
hsTet <- hema %>% select(ID,tet_hs) %>%
  mutate(tet_hs_erg1 = ifelse(tet_hs < 0.01, 0, 1)) %>%
  mutate(tet_hs_erg2 = ifelse(tet_hs <= 0.1, 0, 1)) %>%
  mutate(tet_hs_erg3 = ifelse(tet_hs <= 0.5, 0, 1)) %>%
  mutate(tet_hs_erg4 = ifelse(tet_hs <= 1.0, 0, 1)) %>%
  mutate(tet_hs_erg5 = ifelse(tet_hs <= 5.0, 0, 1))

#create binary variables for Serum results
vTet <- hema %>% select(ID, tet_v) %>%
  mutate(tet_v_erg1 = ifelse(tet_v < 0.01, 0, 1)) %>%
  mutate(tet_v_erg2 = ifelse(tet_v <= 0.1, 0, 1)) %>%
  mutate(tet_v_erg3 = ifelse(tet_v <= 0.5, 0, 1)) %>%
  mutate(tet_v_erg4 = ifelse(tet_v <= 1.0, 0, 1)) %>%
  mutate(tet_v_erg5 = ifelse(tet_v <= 5.0, 0, 1))

#create binary variables for HemaForm results

pTet <- hema %>% select(ID, tet_p) %>%
  mutate(tet_p_erg1 = ifelse(tet_p < 0.01, 0, 1)) %>%
  mutate(tet_p_erg2 = ifelse(tet_p <= 0.1, 0, 1)) %>%
  mutate(tet_p_erg3 = ifelse(tet_p <= 0.5, 0, 1)) %>%
  mutate(tet_p_erg4 = ifelse(tet_p <= 1.0, 0, 1)) %>%
  mutate(tet_p_erg5 = ifelse(tet_p <= 5.0, 0, 1)) %>%
  select(-tet_p)

#Merge two data sets

#HS and VBP
hs_vTet <- merge(hsTet, vTet, by = "ID", all = TRUE) 
hs_vTet <-na.omit(hs_vTet)


#HF and VBP
p_vTet <- merge(pTet, vTet, by = "ID", all = TRUE) 
p_vTet <- na.omit(p_vTet)
 

#Merge HF, VBP, and HS
all_Tet <- merge(pTet,vTet, by ="ID", all=TRUE)
all_Tet <- merge(all_Tet,hsTet,by ="ID", all=TRUE)
all_Tet<- na.omit(all_Tet)

#One data set with only the numeric measures of Anti-Clostridium tetani toxin antibodies
tet_od<-hema %>% select(ID,tet_hs,tet_v,tet_p) %>% 
  na.omit()

tet_od$tet_hs<-as.numeric(tet_od$tet_hs)
tet_od$tet_p<-as.numeric(tet_od$tet_p)
tet_od$tet_v<-as.numeric(tet_od$tet_v)

```
##04: Calculating sensitivites and 95%-CI within the different cut-off values for detection Anti-Clostridium tetani toxin IgG 
according to the manufacture manual of the used ELISA (PTETG043, NovaTec, Germany) [2]
```{r}
#Calculating sensitivity cut off < 0.01
#HS vs VBP
table(all_Tet$tet_hs_erg1,all_Tet$tet_v_erg1)
as.table(matrix(c(49, 0, 0, 0), nrow = 2, byrow = TRUE)) %>% epi.tests()

#HF vs VBP
table(all_Tet$tet_p_erg1,all_Tet$tet_v_erg1)
as.table(matrix(c(49, 0, 0, 0), nrow = 2, byrow = TRUE)) %>% epi.tests()

#Calculating sensitivity cut off <= 0.1
#HS vs VBP
table(all_Tet$tet_hs_erg2,all_Tet$tet_v_erg2)
as.table(matrix(c(49, 0, 0, 0), nrow = 2, byrow = TRUE)) %>% epi.tests()
#HF vs VBP
table(all_Tet$tet_p_erg2,all_Tet$tet_v_erg2)
as.table(matrix(c(49, 0, 0, 0), nrow = 2, byrow = TRUE)) %>% epi.tests()

#Calculating sensitivity cut off <= 0.5
#HS vs VBP
table(all_Tet$tet_hs_erg3,all_Tet$tet_v_erg3)
as.table(matrix(c(46, 0, 3, 0), nrow = 2, byrow = TRUE)) %>% epi.tests()

#HF vs VBP
table(all_Tet$tet_p_erg3,all_Tet$tet_v_erg3)
as.table(matrix(c(49, 0, 0, 0), nrow = 2, byrow = TRUE)) %>% epi.tests()

#Calculating sensitivity cut off <= 1.0
#HS vs VBP
table(all_Tet$tet_hs_erg4,all_Tet$tet_v_erg4)
as.table(matrix(c(33, 0, 14, 2), nrow = 2, byrow = TRUE)) %>% epi.tests()
#HF vs VBP
table(all_Tet$tet_p_erg4,all_Tet$tet_v_erg4)
as.table(matrix(c(40, 0, 7, 2), nrow = 2, byrow = TRUE)) %>% epi.tests()

#Calculating sensitivity cut off <= 5.0 
#HS vs VBP
table(all_Tet$tet_hs_erg5,all_Tet$tet_v_erg5)
as.table(matrix(c(0, 0, 0, 49), nrow = 2, byrow = TRUE)) %>% epi.tests()
#HF vs VBP
table(all_Tet$tet_p_erg5,all_Tet$tet_v_erg5)
as.table(matrix(c(0, 0, 0, 49), nrow = 2, byrow = TRUE)) %>% epi.tests()

```
##05: Correlation of Anti-Clostridium tetani toxin IgG concentration in the different samples. Calculation of Pearson correlation.
```{r}
#Testing normal distribution
hist(tet_od$tet_hs)
hist(tet_od$tet_p)
hist(tet_od$tet_v)

#Testing linearity
ggplot(tet_od, aes(x = tet_hs, y = tet_v)) +
  geom_point(shape = 21, size = 3, fill = rgb(0, 90/255, 160/255), color = "black") + 
  geom_smooth(method = "lm", color = rgb(0, 90/255, 160/255)) + 
  labs(title = "Scatter plot tetanus antibody concentration", subtitle = "r=0.73 (95% CI:0.56-0.83)",
       x = "HemaSpot", y = "venous blood puncture")+ 
  theme(panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank()) 

ggplot(tet_od, aes(x = tet_p, y = tet_v)) +
  geom_point(shape = 21, size = 3, fill = rgb(210/255, 50/255, 100/255), color = "black") +
  geom_smooth(method = "lm", color = rgb(210/255, 50/255, 100/255)) + 
  labs(title = "Scatter plot tetanus antibody concentration", subtitle = "r=0.81 (95% CI:0.69-0.89)",
       x = "HemaForm", y = "venous blood puncture")+ 
  theme(panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank()) 

#Calculation of Pearson correlation

correlation_1<-cor.test(tet_od$tet_hs, tet_od$tet_v, method = "pearson")
correlation_2<-cor.test(tet_od$tet_p, tet_od$tet_v, method = "pearson")

```

##06:Calculation and creation of Bland-Altman plots
```{r}
#Prepare data for Bland-Altman plots

#Plot for comparing HS and VBP
ba1tet <- hema %>% select(tet_v, tet_hs) %>% 
  mutate(tet_v = as.character(tet_v),
        tet_v = as.numeric(tet_v)) %>% 
  mutate(tet_hs = as.character(tet_hs),
         tet_hs = as.numeric(tet_hs)) %>% 
  na.omit()

#Plot for comparing HF and VBP
ba2tet <- hema %>% select(tet_v, tet_p) %>% 
  mutate(tet_v = as.character(tet_v),
         tet_v = as.numeric(tet_v)) %>% 
  mutate(tet_p = as.character(tet_p),
         tet_p = as.numeric(tet_p)) %>% 
  na.omit()

#Plot for comparing HF and HS
ba3tet <- hema %>% select(tet_p, tet_hs) %>% 
  mutate(tet_hs = as.character(tet_hs),
         tet_hs = as.numeric(tet_hs)) %>% 
  mutate(tet_p = as.character(tet_p),
         tet_p = as.numeric(tet_p)) %>% 
  na.omit()

#Plot for comparing HS and VBP
ba1tet_plot <- BlandAltmanLeh::bland.altman.plot(ba1tet$tet_v, ba1tet$tet_hs, graph.sys = "ggplot2") +
  geom_point(color = rgb(0, 90/255, 160/255), size = 3) +
  labs(title = "",
       x = "Average measure of Anti-Clostridium tetani toxin IgG concentration",
       y = "Difference in Anti-Clostridium tetani toxin IgG concentration between sampling methods: 
       HemaSpot-HF - HemaForm") +
  theme_classic()+
  scale_y_continuous(breaks = seq(-0.5, 2.5, 0.5), limits=c(-0.5, 2.5)) +
  scale_x_continuous(breaks = seq(0, 3.0, 0.5), limits=c(0, 3.0)) +
  theme(axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 6),
        panel.border = element_rect(color = "black", fill = NA, size = 1))
  
#Plot for comparing HF and VBP
ba2tet_plot <- BlandAltmanLeh::bland.altman.plot(ba2tet$tet_v, ba2tet$tet_p, graph.sys = "ggplot2") + 
  geom_point(color = rgb(210/255, 50/255, 100/255), size = 3) +
  labs(title = "",
       x = "Average measure of Anti-Clostridium tetani toxin IgG concentration",
       y = "Difference in Anti-Clostridium tetani toxin IgG concentration between sampling methods: 
       HemaSpot-HF - HemaForm") +
  theme_classic()+
  scale_y_continuous(breaks = seq(-0.5, 2.5, 0.5), limits=c(-0.5, 2.5)) +
  scale_x_continuous(breaks = seq(0, 3.0, 0.5), limits=c(0, 3.0)) +
  theme(axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 6),
        panel.border = element_rect(color = "black", fill = NA, size = 1))

#Plot for comparing HF and HS
ba3tet_plot <- BlandAltmanLeh::bland.altman.plot(ba3tet$tet_p, ba3tet$tet_hs, graph.sys = "ggplot2") + 
  geom_point(color = "forestgreen", size = 3) +
  labs(title = "",
       x = "Average measure of Anti-Clostridium tetani toxin IgG concentration",
       y = "Difference in Anti-Clostridium tetani toxin IgG concentration between sampling methods: 
       HemaSpot-HF - HemaForm") +
  theme_classic()+
  scale_y_continuous(breaks = seq(-0.5, 2.5, 0.5), limits=c(-0.5, 1.0)) +
  scale_x_continuous(breaks = seq(0, 3.0, 0.5), limits=c(0, 3.0)) +
  theme(axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 6),
        panel.border = element_rect(color = "black", fill = NA, size = 1))


```