---
title: "02_code_questionnaire_analysis_Kettlitz et al."
author: "Robyn Kettlitz"
date: "2024-11-17"
output: html_document
---

R code of the article: 
Kettlitz et al., Feasibility and diagnostic performance of blood self-sampling with HemaSpot HF: The use case of Anti-Clostridium tetani toxin IgG detection.

This script was used to analyse the feasibility including acceptance of DBS self-sampling with HS. 

For a description of variables please refer to the codebook.

Abbreviation: 

HS = HemaSpot
HF = HemaForm
VBP = venous blood puncture

References: 
[1]: Sciences SO. HemaSpot™-HF Blood Collection Device  [Available from: https://www.spotonsciences.com/hemaspot-hf/ ]
[2]: NovaTec’s Clostridium tetani toxin 5S IgG plus – ELISA (PTETG043, NovaTec, Germany)

Use of artificial intelligence:
We used OpenAI's ChatGPT v2 to review and refine the R code for improved structure and functionality 

Code review by Ruibing Shi [Department for Epidemiology, Helmholtz Centre for Infection Research, Braunschweig, Lower Saxony, Germany]
-----------------------------
Codebook:


Variable name| Content question| Answer options

tet_hs | Result U/ml tetanus HS | continuous
tet_v | Result U/ml tetanus VBP | continuous
tet_p | Result U/ml tetanus HF  | continuous

hs1 | Recieved HS device | 0: no, 1: yes
auf0 | HS without conspicuous features | 0: no, 1: yes
auf1 | HS with external conspicuous features | 0: no, 1: yes
auf2 | HS with internal conspicuous features | 0: no, 1: yes
besch1 | HS external damage | 0: no, 1: yes
			
sex| gender| female:0, male:1, other:2, missing: 3
age |	age |	< 30:0, 30-49:1, > 50: 2, missing:3
edu	| education - does your highest school-leaving qualification correspond to the Abitur or subject-specific Abitur?| 	no:0, yes:1, I don't know:2, missing:	3
exp_0 | Have you ever drawn blood before this study? | no:0, yes:1, missing:	2
exp_1 |	Blood collection from others| | no:0, yes:1 , missing:	2
exp_2	|	Blood self-sampling | no:0, yes:1, missing:	2
			
meth_0| Which method(s) did you use to improve blood circulation prior to self-blood sampling? | no:0, yes:1, missing	3
meth_1|		Hold hand/fingers under very warm water | no:0, yes:1, missing	3
meth_2|		Shake hand vigorously | no:0, yes:1, missing	3
meth_3|		Massage/rub fingers | no:0, yes:1, missing	3
meth_4|		"Pump" (open and close fist) | no:0, yes:1, missing	3
meth_5|		Move arm quickly back and forth while hanging down| no:0, yes:1, missing	3
			
toL |	How many of the enclosed lancing devices (lancets) did you use?	| Do not click:	0, one:	1, two:	2, missing: 3
erst |	Were you able to wipe away the first drop of blood?|	no:0, yes:1, don't know:	2,missing:	3
tropf|	How many drops of blood did you need to wet the filter in the blue click box? |	 I couldn't tell the difference between the drops of blood:	0, 1:1, 2-3:2,, more than 3:3, missing:	4, SPECIAL CASE: more than 3 + drops of blood cannot be distinguished from each other:5
			
nach | Has the puncture site bled longer and/or heavily?| no:0, yes:1, missing:	3
ueb | How much effort did it take you to draw your own blood?| None at all:0, something:1, much:2, Very much:3, missing:4
sprach |	Which instructions did you mainly use? | Original instructions (leaflet):0, German study guide (booklet):	1,missing:2
verst |	To what extent were you able to understand these instructions?|	Easy to understand:0, Partly understandable:1, Difficult to understand:2,missing:3
insg|	On a scale from 0 (very negative) to 10 (very positive), how was your overall experience of self-blood sampling?	0 - very negative:0 to
		10 - very positive:10 ; missing:	11
			
akz | The performance of the blood self-sampling was acceptable. I agree ...|	not at all.:0, rather not:1, rather too.:2, fully and completely.:3, missing:4
			
benu |	The blood self-sampling procedure was user-friendly. I agree ...| not at all.:0, rather not:1, rather too.:2, fully and completely.:3, missing:4
			
schw| The total blood self-sampling was ... very difficult:	0, difficult:	1, simple:	2, very simple:	3, missing:	4
			
stich | The use of the stitching device was ... very difficult:	0, difficult:	1, simple:	2, very simple:	3, missing:	4
			
aufbr| Applying the blood to the filter was...	very difficult:	0, difficult:	1, simple:	2, very simple:	3, missing:	4
			
schm | on a scale from 0 (no pain at all) to 10 (worst pain imaginable), how much pain did the self-drawing of blood cause?	0 - no pain at all to
		10 - strongest imaginable pain, missing:11
			
zum |	Was this pain reasonable for you in the context of this study?|	Yes	0, partly: 1, no: 2, missing:	3
			
praef | Assuming that the laboratory result of self-sampling is the same as that of venous blood sampling by medical staff, which method would you prefer and why?|	Blood self-sampling (on the finger):0, Venous blood sampling by a doctor (on the arm):1, I don't know:2, missing:3, Invalid:4
			
selbst | Blood self-sampling (on the finger), because ... |	
    Invalid - more than 1 ticked:	0,
		□ less time is required (travel, appointments/waiting time):	1,
		□ I can determine the time and place myself:	2,
		□ I don't like going to the doctor: 	3,
		□ I like to act on my own responsibility:  	4,
		□ it gives me a sense of security to do things myself:	5,  
		□ it is simple and uncomplicated:	6,
		□ I make taking blood samples less painful: 	7,
		□ the risk of infection is lower:	8,
		Missing:	9
			
			
med	| Venous blood sampling by medical personnel (on the arm) because ...|
    Invalid:	0,
		□ less time is required (blood sampling):	1,
		□ I don't want to hurt myself: 	2,
		□ I am afraid of doing something wrong:	3,
		□ it gives me a feeling of security when it is done by specialist staff:	4,
		□ it is simple and uncomplicated:	5,
		□ blood sampling is less painful due to trained personnel: 	6,
		□ the risk of infection is lower:	7,
		Missing	8
			
empf | Would you recommend participation in a study with blood self-sampling to your friends?|	no:0, yes:1, possibly:2, missing:	3
			
-----------------------------

##00: Set up, loading relevant packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(pacman)

pacman::p_load(tidyverse, haven, tableone, broom, lmtest,car, plotly,gridExtra, patchwork,ggpubr,cowplot,gtsummary,gt, summarytools)

```

##01: Read data
```{r}
#identify newest file
recent_file <- list.files(path = "../../../5_Daten/Labordaten/Bearbeitet/", pattern = "MergedAndCleanedData_jor", full.names = TRUE) %>% sort(decreasing = TRUE) %>% 
  .[1]

hema <- read_csv(recent_file) %>% 
  select(-1)
```
##02: Counts of delivered questionnaires by HS and VBP
```{r}
#Counts....

#Overall questionnaires for analysis 

#Instruction questionnaire
hema %>% filter(anl1==1) %>% 
  count()
#Acceptance questionnaire
hema %>% filter(fb1==1) %>% 
  count()
#Feedback questionnaire
hema %>% filter(!empf==3) %>% filter(!is.na(empf)) %>% 
  count()

#Stratified by groups (with and without VBP)
#Only HS without additional VBP
hema %>% filter (!is.na(hs1==1) &is.na(od2_v)) %>% count()

#Acceptance questionnaire of HS without VBP
hema %>% filter (!is.na(hs1==1) &is.na(od2_v)& fb1==1) %>% count()

#Instruction questionnaire without additional VBP
hema %>% filter (!is.na(hs1==1) &is.na(od2_v)& anl1==1) %>% count()


#HS and VBP for analysis of questionnaires 
hema %>% filter(!is.na(od2_v) & hs1==1) %>% 
  count()

#Instruction questionnaire with VBP
hema %>% filter(!is.na(od2_v) & anl1==1 & hs1==1) %>% 
  count()

#Acceptance questionnaire with VBP
hema %>% filter(!is.na(od2_v) & fb1==1 & hs1==1) %>%
  count()

```

##03: Further data preparation for summary statistics (Rename the hema_plot)
```{r}
#generate variable akzlog for a binary display of accpetance (to stratify data by this variable)
hema <- hema %>%
  mutate(akzlog = case_when(akz == 0 ~ 0,
                            akz == 1 ~ 0,
                            akz == 2 ~ 1,
                            akz == 3 ~ 1))

#Frequency table with the accepatnce question
#the percentages were calculated and added to the Word Table
#Additional visulaitation:

hema_data1<- hema %>% 
  mutate(akz=case_when(akz==0~"Not at all",
                       akz==1~"Rather not",
                       akz==2~"Rather",
                       akz==3~"Fully"),
         benu=case_when(benu==0~"Not at all",
                       benu==1~"Rather not",
                       benu==2~"Rather",
                       benu==3~"Fully"),
          age=case_when(age==0~"<30",
                        age==1~"30-49",
                        age==2 ~"≥50"),
          edu=case_when(edu==0 ~ "Abitur: no",
                        edu==1 ~ "Abitur: yes",
                        edu==2 ~ "Don't know"),
          exp_0=case_when(exp_0==0~"No experience with blood sampling",
                          exp_0==1~"Yes experience with blood sampling"),
         sex=case_when(sex==0~"Female",
                       sex==1~"Male"),
         ueb=case_when(ueb==0~ "No effort at all", 
                      ueb==1~"A bit effort", 
                      ueb==2~"Much effort", 
                      ueb==3~"Very much effort"),
         praef=case_when(praef==0~ "Blood self-sampling", 
                         praef==1~ "Venous blood puncture by medical staff", 
                         praef== 2~"Don't know" ),
         empf=case_when(empf==0~"No",
                        empf==1~"Yes",
                        empf==2~ "Maybe recommend"),
         meth_0==case_when(meth_0==0~"No, method used for blood flow",
                           meth_0==1~ "Yes, method used for blood flow"),
         anzL==case_when(anzL==1~"One lancet used",
                         anzL==2~"Two lancet used"),
         erst==case_when(erst==0~ "Not removed first drop",
                         erst==1~ "Yes removed first drop"),
         tropf=case_when(tropf==0~"Don't know",
                         tropf==1~"1",
                         tropf==2~"2-3",
                         tropf==3|tropf==5~"3+"),
         verst=case_when(verst==0~"Easy understandable",
                         verst==1~ "Partially understandable",
                         verst==2~ "Hard to understand"),
         zum=case_when(zum==0~"Yes,acceptable",
                       zum==1~"Yes,partially acceptable",
                       zum==2~"No,not acceptable"),
         nach=case_when(nach==0~"No",
                        nach==1~"Yes"),
         aufbr=case_when(aufbr==0 ~"Very difficult",
                         aufbr==1 ~"Difficult",
                         aufbr==2~"Easy",
                         aufbr==3 ~"Very easy"),
         stich=case_when(stich==0 ~"Very difficult",
                         stich==1 ~"Difficult",
                         stich==2~"Easy",
                         stich==3 ~"Very easy"),
         schw=case_when(schw==0 ~"Very difficult",
                         schw==1 ~"Difficult",
                         schw==2~"Easy",
                         schw==3 ~"Very easy"),
         selbst=case_when(selbst==1~"less time is needed (travel, appointments/waiting time)",
                          selbst==2~ "I can determine the time and place myself",
                          selbst==3~"I don't like going to the doctor",
                          selbst==4~"I like to take responsibility for my own care ",
                          selbst==5~"It makes me feel safe to do things myself",
                          selbst==6~"It is simple and straightforward",
                          selbst==7~"I make the blood sample less painful",
                          selbst==8~"The risk of infection is lower"),
          med=case_when(med==1~"It takes less time (blood sample)",
                          med==2~ "I do not want to hurt myself",
                          med==3~"I am afraid of doing something wrong",
                          med==4~"It gives me a sense of security when it is done by trained personnel",
                          med==5~"It is simple and uncomplicated",
                          med==6~"When it is done by trained personnel, the blood sample causes less pain",
                          med==7~"The risk of infection is lower"))


#Adjust the NAs in the data
hema_data1$age[hema_data1$age == 3] <- NA
hema_data1$edu[hema_data1$edu == 3] <- NA
hema_data1$exp_0[hema_data1$exp_0 == 2] <- NA
hema_data1$sex[hema_data1$sex == 2] <- NA
hema_data1$akz[hema_data1$akz == 4] <- NA
hema_data1$ueb[hema_data1$ueb == 4] <- NA
hema_data1$praef[hema_data1$praef == 4] <- NA
hema_data1$praef[hema_data1$praef == 3] <- NA
hema_data1$empf[hema_data1$empf == 3] <- NA
hema_data1$schm[hema_data1$schm == 11] <- NA
hema_data1$meth_0[hema_data1$meth_0 == 3] <- NA
hema_data1$meth_1[hema_data1$meth_1 == 3]<- NA
hema_data1$meth_2[hema_data1$meth_2 == 3]<- NA
hema_data1$meth_3[hema_data1$meth_3 == 3]<- NA
hema_data1$meth_4[hema_data1$meth_4 == 3]<- NA
hema_data1$meth_5[hema_data1$meth_5 == 3]<- NA
hema_data1$anzL[hema_data1$anzL == 3]<- NA
hema_data1$erst[hema_data1$erst == 3]<- NA
hema_data1$tropf[hema_data1$tropf == 4]<- NA
hema_data1$verst[hema_data1$verst == 3]<- NA
hema_data1$zum[hema_data1$zum == 3]<- NA
hema_data1$nach[hema_data1$nach == 3]<- NA
hema_data1$aufbr[hema_data1$aufbr == 4]<- NA
hema_data1$selbst[hema_data1$selbst == 9]<- NA
hema_data1$selbst[hema_data1$selbst == 0]<- NA
hema_data1$med[hema_data1$med == 9]<- NA
hema_data1$med[hema_data1$med == 0]<- NA
hema_data1$schw[hema_data1$schw == 4]<- NA
 

#Save as numeric variables
hema_data1 <- hema_data1%>%
  mutate(insg = as.numeric(insg))

hema_data1 <- hema_data1%>%
  mutate(schm = as.numeric(as.character(schm)))
```
##03: Create a summary tables with overall column and stratification for supplement
```{r}
#Overall sample arrived HemaSpot

hema_data1 %>%
  select(age, sex, edu, exp_0) %>%
  tbl_summary( # Variables to summarize
    missing = "always",
    statistic = list(
    all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
    all_categorical() ~ "{n} ({p}%)" ), 
    # Count and percentage for categorical variables
    # Show all missing values
  ) 

#Only those with additional VBP

hema_data1 %>%
  filter(!is.na(tet_hs), !is.na(tet_v)) %>% 
  select(age, sex, edu, exp_0) %>%
  tbl_summary( # Variables to summarize
    missing = "always",
    statistic = list(
    all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
    all_categorical() ~ "{n} ({p}%)" ) # Count and percentage for categorical variables
    # Show all missing values
  ) 
```

##03a: Cacluation of percentange and signfificance tests (group comparison) for Figure 2: Study participants’ characteristics (shown in percentage) of self-blood. For the question: Would you recommend participation in a study for blood self-sampling to your friends? Stratified by experience with blood-sampling, age, sex, and education 
```{r}
hema_data1 %>% 
  select(empf) %>% 
  group_by(empf) %>% # Group by exp_0
  summarise(count = sum(!is.na(empf))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places

hema_data1 %>% 
  select(exp_0, empf) %>% 
  group_by(exp_0, empf) %>% # Group by exp_0
  summarise(count = sum(!is.na(empf))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places


hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(edu, empf) %>% # Group by Alter
  summarise(count = sum(!is.na(empf))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(sex, empf) %>% # Group by Alter
  summarise(count = sum(!is.na(empf))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(age, empf) %>% # Group by Alter
  summarise(count = sum(!is.na(empf))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 



hema_data1 %>%
  # Remove rows with missing values in empf and age, sex, edu, exp_0
  filter(!is.na(empf) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  
  # Perform Chi-square tests
  summarise(
    chi_sq_empf_age = chisq.test(table(empf, age))$p.value,
    chi_sq_empf_sex = chisq.test(table(empf, sex))$p.value,
    chi_sq_empf_edu = chisq.test(table(empf, edu))$p.value,
    chi_sq_empf_exp_0 = chisq.test(table(empf, exp_0))$p.value
  )

#Use Fisher's Exact Test: 
hema_data1 %>%
  filter(!is.na(empf) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  summarise(
    fisher_empf_age = fisher.test(table(empf, age))$p.value,
    fisher_empf_sex = fisher.test(table(empf, sex))$p.value,
    fisher_empf_edu = fisher.test(table(empf, edu))$p.value,
    fisher_empf_exp_0 = fisher.test(table(empf, exp_0))$p.value
  )

```

##03b: Cacluation of percentange and significance tests (group comparison)  for Figure 2: Study participants’ characteristics (shown in percentage) of self-blood. For the question:To what extent do you agree with the following statements? The procedure of blood self-sampling was acceptable. Stratified by experience with blood-sampling, age, sex, and education 
```{r}
hema_data1 %>% 
  select(akz) %>% 
  group_by(akz) %>%  # Group by exp_0
  summarise(count = sum(!is.na(akz))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places

hema_data1 %>% 
  select(exp_0, akz) %>% 
  group_by(exp_0, akz) %>% # Group by exp_0
  summarise(count = sum(!is.na(akz))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places


hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(edu, akz) %>% # Group by Alter
  summarise(count = sum(!is.na(akz))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(sex, akz) %>% # Group by Alter
  summarise(count = sum(!is.na(akz))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(age, akz) %>% # Group by Alter
  summarise(count = sum(!is.na(akz))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>%
  # Remove rows with missing values in akz and age, sex, edu, exp_0
  filter(!is.na(akz) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  
#Perform Chi-square tests
  summarise(
    chi_sq_akz_age = chisq.test(table(akz, age))$p.value,
    chi_sq_akz_sex = chisq.test(table(akz, sex))$p.value,
    chi_sq_akz_edu = chisq.test(table(akz, edu))$p.value,
    chi_sq_akz_exp_0 = chisq.test(table(akz, exp_0))$p.value
  )

#Use Fisher's Exact Test: 
hema_data1 %>%
  filter(!is.na(akz) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  summarise(
    fisher_akz_age = fisher.test(table(akz, age))$p.value,
    fisher_akz_sex = fisher.test(table(akz, sex))$p.value,
    fisher_akz_edu = fisher.test(table(akz, edu))$p.value,
    fisher_akz_exp_0 = fisher.test(table(akz, exp_0))$p.value
  )

```
##03c:Cacluation of percentange and significance tests (group comparison) for Figure 2: Study participants’ characteristics (shown in percentage) of self-blood. For the question: How much effort did it imply for you, to draw blood self-sampling? Stratified by experience with blood-sampling, age, sex, and education 
```{r}
hema_data1 %>% 
  select(ueb) %>% 
  group_by(ueb) %>%  # Group by exp_0
  summarise(count = sum(!is.na(ueb))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places

hema_data1 %>% 
  select(exp_0, ueb) %>% 
  group_by(exp_0, ueb) %>% # Group by exp_0
  summarise(count = sum(!is.na(ueb))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places


hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(edu, ueb) %>% # Group by Alter
  summarise(count = sum(!is.na(ueb))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(sex, ueb) %>% # Group by Alter
  summarise(count = sum(!is.na(ueb))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(age, ueb) %>% # Group by Alter
  summarise(count = sum(!is.na(ueb))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 


hema_data1 %>%
  # Remove rows with missing values in ueb and age, sex, edu, exp_0
  filter(!is.na(ueb) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  
  # Perform Chi-square tests
  summarise(
    chi_sq_ueb_age = chisq.test(table(ueb, age))$p.value,
    chi_sq_ueb_sex = chisq.test(table(ueb, sex))$p.value,
    chi_sq_ueb_edu = chisq.test(table(ueb, edu))$p.value,
    chi_sq_ueb_exp_0 = chisq.test(table(ueb, exp_0))$p.value
  )

#Use Fisher's Exact Test: 
hema_data1 %>%
  filter(!is.na(ueb) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  summarise(
    fisher_ueb_age = fisher.test(table(ueb, age))$p.value,
    fisher_ueb_sex = fisher.test(table(ueb, sex))$p.value,
    fisher_ueb_edu = fisher.test(table(ueb, edu))$p.value,
    fisher_ueb_exp_0 = fisher.test(table(ueb, exp_0))$p.value
  )

```

##03d: Cacluation of percentange and significance tests (group comparison) for Figure 2: Study participants’ characteristics (shown in percentage) of self-blood. For the question: How much effort did it imply for you, to draw blood self-sampling? Stratified by experience with blood-sampling, age, sex, and education 
```{r}
hema_data1 %>% 
  select(praef) %>% 
  group_by(praef) %>% # Group by exp_0
  summarise(count = sum(!is.na(praef))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places

hema_data1 %>% 
  select(exp_0, praef) %>% 
  group_by(exp_0, praef) %>% # Group by exp_0
  summarise(count = sum(!is.na(praef))) %>%# Count the non-missing entries per exp_0-Gruppe
  mutate(percentage = round(count / sum(count) * 100, 0))  # Percentage per exp_0rounded to 0 decimal places


hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(edu, praef) %>% # Group by Alter
  summarise(count = sum(!is.na(praef))) %>%# Count the non-missing entries per  group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(sex, praef) %>% # Group by Alter
  summarise(count = sum(!is.na(praef))) %>%# Count the non-missing entries per group
  mutate(percentage = round(count / sum(count) * 100, 0)) 

hema_data1 %>% 
  select(age, sex, edu, exp_0, praef, empf, ueb, akz) %>% 
  group_by(age, praef) %>% # Group by Alter
  summarise(count = sum(!is.na(praef))) %>%# Count the non-missing entries per group
  mutate(percentage = round(count / sum(count) * 100, 0)) 



hema_data1 %>%
  # Remove rows with missing values in praef and age, sex, edu, exp_0
  filter(!is.na(praef) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  
  # Perform Chi-square tests
  summarise(
    chi_sq_praef_age = chisq.test(table(praef, age))$p.value,
    chi_sq_praef_sex = chisq.test(table(praef, sex))$p.value,
    chi_sq_praef_edu = chisq.test(table(praef, edu))$p.value,
    chi_sq_praef_exp_0 = chisq.test(table(praef, exp_0))$p.value
  )

#Use Fisher's Exact Test: 
hema_data1 %>%
  filter(!is.na(praef) & !is.na(age) & !is.na(sex) & !is.na(edu) & !is.na(exp_0)) %>%
  summarise(
    fisher_praef_age = fisher.test(table(praef, age))$p.value,
    fisher_praef_sex = fisher.test(table(praef, sex))$p.value,
    fisher_praef_edu = fisher.test(table(praef, edu))$p.value,
    fisher_praef_exp_0 = fisher.test(table(praef, exp_0))$p.value
  )
```
##04: Create a summary tables with overall column and stratification for supplement by row
```{r}

hema_data1 %>%
  select(age, sex, edu, exp_0, akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach, stich,anzL, schw) %>%
  tbl_summary(
    by = age,  # Stratify by age
    include = c(akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach, stich,anzL, schw),  # Variables to summarize
    missing = "always",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
      all_categorical() ~ "{n} ({p}%)"  # Count and percentage for categorical variables
    ), percent="row",
    type = list(schm ~ "continuous"), # Force 'schm' to be treated as continuous
  ) %>%
  add_overall()

hema_data1 %>%
  select(age, sex, edu, exp_0, akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach, stich,anzL, schw) %>%
  tbl_summary(
    by = sex,  # Stratify by age
    include = c(akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach,stich,anzL, schw),  # Variables to summarize
    missing = "always",
   statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
      all_categorical() ~ "{n} ({p}%)"  # Count and percentage for categorical variables
    ),
    type = list(schm ~ "continuous"), percent = "row",  # Force 'schm' to be treated as continuous
  ) %>%
  add_overall()

hema_data1 %>%
  select(age, sex, edu, exp_0, akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach, stich,anzL, schw) %>%
  tbl_summary(
    by = edu,  # Stratify by age
    include = c(akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach, stich,anzL, schw),  # Variables to summarize
    missing = "always",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
      all_categorical() ~ "{n} ({p}%)"  # Count and percentage for categorical variables
    ),
    type = list(schm ~ "continuous"), percent = "row",  # Force 'schm' to be treated as continuous
  ) %>%
  add_overall()


hema_data1 %>%
  select(age, sex, edu, exp_0, akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach, stich,anzL, schw) %>%
  tbl_summary(
    by = exp_0,  # Stratify by age
    include = c(akz,benu,insg, ueb,aufbr,praef, selbst, med, empf,zum, schm ,meth_0, meth_1, meth_2, meth_3, meth_4, meth_5, erst, tropf, verst, nach, stich,anzL, schw),  # Variables to summarize
    missing = "always",
   statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
      all_categorical() ~ "{n} ({p}%)"  # Count and percentage for categorical variables
    ),
    type = list(schm ~ "continuous"), percent = "row",  # Force 'schm' to be treated as continuous
  ) %>%
  add_overall()


```
##05: Net Promoter Score calulations ad stratified by experience with blood-sampling, age, sex, and education 
```{r}
df <- hema %>%
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps <- df[2,2]/(df[2,2]+df[1,2]) - df[1,2]/(df[2,2]+df[1,2]) 
#calculate Net Promoter Score with resulting table df (df[2,2] represents the count of those who agree| df[1,2] represents the count of those who disagree) ->#(106/(106+3)) - (3/(106+3))

nps


#Net Promoter Score calulations by male
table(hema$sex)

df_sex_f <- hema %>% 
  filter (sex==0) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_sex_f <- df_sex_f[2,2]/(df_sex_f[2,2]+df_sex_f[1,2]) - df_sex_f[1,2]/(df_sex_f[2,2]+df_sex_f[1,2])#Female 97% 
#calculate Net Promoter Score with resulting table df (df[2,2] represents the count of those who agree| df[1,2] represents the count of those who disagree) ->#(106/(106+3)) - (3/(106+3))

df_sex_m <- hema %>% 
  filter (sex==1) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_sex_m <- df_sex_m[2,2]/(df_sex_m[2,2]+df_sex_m[1,2]) - df_sex_m[1,2]/(df_sex_m[2,2]+df_sex_m[1,2])#Male: 91%


#Net Promoter Score calulations by <30

df_age_0 <- hema %>% 
  filter (age==0) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_age_0<- df_age_0[2,2]/(df_age_0[2,2]+df_age_0[1,2]) - df_age_0[1,2]/(df_age_0[2,2]+df_age_0[1,2])#82%


#Net Promoter Score calulations by 30-49

df_age_1 <- hema %>% 
  filter (age==1) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_age_1<- df_age_1[2,2]/(df_age_1[2,2]+df_age_1[1,2]) - df_age_1[1,2]/(df_age_1[2,2]+df_age_1[1,2])


#Net Promoter Score calulations by >50

df_age_2 <- hema %>% 
  filter (age==2) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_age_2<- df_age_2[2,2]/(df_age_2[2,2]+df_age_2[1,2]) - df_age_2[1,2]/(df_age_2[2,2]+df_age_2[1,2])

#Net Promoter Score calulations by education

table(hema$edu)

df_edu_0 <- hema %>% 
  filter (edu==0) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_edu_0<- df_edu_0[2,2]/(df_edu_0[2,2]+df_edu_0[1,2]) - df_edu_0[1,2]/(df_edu_0[2,2]+df_edu_0[1,2])


df_edu_1 <- hema %>% 
  filter (edu==1) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_edu_1<- df_edu_1[2,2]/(df_edu_1[2,2]+df_edu_1[1,2]) - df_edu_1[1,2]/(df_edu_1[2,2]+df_edu_1[1,2])#93%

df_edu_2 <- hema %>% 
  filter (edu==2) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_edu_2<- df_edu_2[2,2]/(df_edu_2[2,2]+df_edu_2[1,2]) - df_edu_2[1,2]/(df_edu_2[2,2]+df_edu_2[1,2])#93%

#Net Promoter Score calulations by experience with blood sampling
table(hema$exp_0)

df_exp_0 <- hema %>% 
  filter (exp_0==0) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_exp_0 <- df_exp_0[2,2]/(df_exp_0[2,2]+df_exp_0[1,2]) - df_exp_0[1,2]/(df_exp_0[2,2]+df_exp_0[1,2])#97%

df_exp_1 <- hema %>% 
  filter (exp_0==1) %>% 
  select(empf) %>%
  filter(empf <= 1) %>% #only include the ones agreeing or disagreeing (so 0 and 1)
  count(empf)
  
nps_exp_1 <- df_exp_1[2,2]/(df_exp_1[2,2]+df_exp_1[1,2]) - df_exp_1[1,2]/(df_exp_1[2,2]+df_exp_1[1,2])#92%


```