---
title: "03_code_non-responder_questionnaire_analysis_Kettlitz et al."
author: "Robyn Kettlitz"
date: "2024-11-17"
output: html_document
---

R code of the article: 
Kettlitz et al., Feasibility and diagnostic performance of blood self-sampling with HemaSpot HF: The use case of Anti-Clostridium tetani toxin IgG detection.

This script was used to analyse the non-responder questionnaire. 
For a description of variables please refer to the codebook.

Abbreviation: 

HS = HemaSpot
HF = HemaForm
VBP = venous blood puncture

References: 
[1]: Sciences SO. HemaSpot™-HF Blood Collection Device  [Available from: https://www.spotonsciences.com/hemaspot-hf/ ]

Use of artificial intelligence:
We used OpenAI's ChatGPT v2 to review and refine the R code for improved structure and functionality 

Code review by Ruibing Shi [Department for Epidemiology, Helmholtz Centre for Infection Research, Braunschweig, Lower Saxony, Germany]
-------------------------------------------
Codebook: 

Question: Indicate why you decided to not participate in the study for blood self-sampling? 

Multiple answers possible: 

Answers: 

F1_1: I have physcial disabilities/ health restrictions that do nor allow me to participate 
F1_2: Participating is too time-consuming
F1_3: I generally have inhibitions regarding drawing of blood samples.
F1_4: I am afraid of blood self-sampling.
F1_5:I don't see the needs annd benefits of the study. 
F1_6: I don't find the instrument for blood self-sampling very trustworthy.
F1_7: I have concerns about data protection
F1_8: I generally don't participate in surveys or studies
F1_9: I generally don't like to participate in studies, that are carried out at my workplace.#
F1_10: Another reason, namely: 
F1_10a: show answers listed for "other reason"
---------------------------------------------

##00: Set up, loading relevant packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(pacman)

pacman::p_load(tidyverse, readstata13, janitor, gtsummary,gt, summarytools)


```

##01: Read data

```{r}
nonr <- read.dta13("S:\\PROJACTIVE\\HEMASPOT-1\\5_Daten\\EpiInfo\\Datenabgleich\\Nonresponder\\2_Datenabgleich Sonja und Julia\\FinalVersion_jor_190617.dta") %>% 
  as_tibble()
```

##02: Descriptive statistics of the answered questions of the non-repsonder questionnaire
```{r}
#count answers
nonr %>% count(F1_1)#I have physcial disabilities/ health restrictions that do nor allow me to participate 
nonr %>% count(F1_2)# Participating is too time-consuming
nonr %>% count(F1_3)# I generally have inhibitions regarding drawing of blood samples.
nonr %>% count(F1_4)# I am afraid of blood self-sampling.
nonr %>% count(F1_5)#I don't see the needs annd benefits of the study. 
nonr %>% count(F1_6)# I don't find the instrument for blood self-sampling very trustworthy.
nonr %>% count(F1_7)# I have concerns about data protection
nonr %>% count(F1_8)# I generally don't participate in surveys or studies
nonr %>% count(F1_9)# I generally don't like to participate in studies, that are carried out at my workplace.
nonr %>% count(F1_10) #Another reason, namely: 
nonr %>% count(F1_10a) #show answers listed for "other reason"
```
#03: 
```{r}
nonr<-nonr %>% mutate(age=case_when(age==0~"<30",
                        age==1~"30-49",
                        age==2 ~"≥50"),
sex=case_when(sex==0~"Female",
                       sex==1~"Male"),
edu=case_when(edu==0 ~ "Abitur: no",
                        edu==1 ~ "Abitur: yes",
                        edu==2 ~ "Don't know"))

nonr$age[nonr$age == 3] <- NA
nonr$edu[nonr$edu == 3] <- NA
nonr$sex[nonr$sex == 2] <- NA


nonr%>%
  select(age, sex, edu, F1_1,F1_2,F1_3,F1_4,F1_5,F1_6,F1_7,F1_8,F1_9,F1_10) %>%
  tbl_summary(
    by = age,  # Stratify by age
    include = c(F1_1,F1_2,F1_3,F1_4,F1_5,F1_6,F1_7,F1_8,F1_9,F1_10),  # Variables to summarize
    missing = "always",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
      all_categorical() ~ "{n} ({p}%)"  # Count and percentage for categorical variables
    ),  percent = "row" ,  # Force 'schm' to be treated as continuous
  ) %>%
  add_overall()

nonr%>%
  select(age, sex, edu, F1_1,F1_2,F1_3,F1_4,F1_5,F1_6,F1_7,F1_8,F1_9,F1_10) %>%
  tbl_summary(
    by = sex,  # Stratify by age
    include = c(F1_1,F1_2,F1_3,F1_4,F1_5,F1_6,F1_7,F1_8,F1_9,F1_10),  # Variables to summarize
    missing = "always",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
      all_categorical() ~ "{n} ({p}%)"  # Count and percentage for categorical variables
    ),percent = "row" , # Force 'schm' to be treated as continuous
  ) %>%
  add_overall()

nonr%>%
  select(age, sex, edu, F1_1,F1_2,F1_3,F1_4,F1_5,F1_6,F1_7,F1_8,F1_9,F1_10) %>%
  tbl_summary(
    by = edu,  # Stratify by age
    include = c(F1_1,F1_2,F1_3,F1_4,F1_5,F1_6,F1_7,F1_8,F1_9,F1_10),  # Variables to summarize
    missing = "always",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",  # Median with IQR for numeric variables
      all_categorical() ~ "{n} ({p}%)"  # Count and percentage for categorical variables
    ),percent = "row" , # Force 'schm' to be treated as continuous
  ) %>%
  add_overall()
```

